Building a continuous benchmarking ecosystem in bioinformatics

### 2024-09-23 manifesto-main.tex
more

\begin{abstract}
Benchmarking, which involves collecting reference datasets and demonstrating method performances, is a requirement for the development of new computational tools, but also becomes a domain of its own to achieve neutral comparisons of methods. Although a lot has been written about how to design and conduct benchmark studies, this perspective sheds light on a wish list for a computational platform to orchestrate benchmark studies. We discuss various ideas for organizing reproducible software environments, formally defining benchmarks, orchestrating standardized workflows, and how they interface with computing infrastructure.
\end{abstract}

@@

In this work, we give our perspective on how to build a high-standard benchmarking ecosystem in the field of bioinformatics, coming primarily from the viewpoint of methodologists (i.e., researchers with computational background working on tools for data). In particular, our case is one of computational scientists embedded in data-rich biology (i.e., computational biology and bioinformatics), where rapid technology development is a distinguishing factor (e.g., single-cell assays coupled to high-throughput sequencing, imaging, etc.); this paucity in turn results in a rapid response from the computational community to develop tools, models, and software for the analysis of such emerging datasets. Although our viewpoint is `bioinformatics', there are analogs in adjacent disciplines where methodologists are creating computational tools in a certain scientific context. One critical step in method development is benchmarking (further defined below), whereby either new methods are compared against existing ones, or a set of existing methods are compared in a more neutral way; for these, we will make use of the terms (and abbreviations) `methods-development papers' (MDPs) and `benchmark-only papers' (BOPs), respectively, defined in a recent meta-analysis study \cite{Cao2023-jz}. Notably, there are various ways to structure the running of a benchmarking study: it can be run by a single person or a small group on a laptop (or a server), or it can be organized as a `challenge' or within the duration of a hackathon (groups of people that come together to competitively assess their approaches and/or populate benchmark components).

Our goal is to lay out general benchmark parameters and definitions,  while abstracting towards an ideal \textit{benchmarking system}. This ranges from general benchmark conceptualization to various technical specifications. We will offer some remarks on the philosophy around the benchmarking activity (e.g., gatekeeping), as well as some technical recommendations (e.g., standards).

@@

Benchmark artifacts can be seen as a collection of data files and (perhaps diverse) source code implementations assembled in a pipeline topology (e.g., methods are run on some input data, and method outputs act as input for metrics) that were run as a workflow in some computing environment. Current practices in single-cell RNA-seq benchmarking show that code is frequently open, reusable and versioned; input datasets are also typically available \cite{Sonrel2023-te}. However, the \textit{extensibility} of benchmarking studies is generally very low, as well as the proportion of benchmarks using a formal workflow system. Thus, a formal system to organize components and orchestrate the crossing of datasets, methods, and metrics would largely impact the current landscape.

@@

Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning or being aware of homogeneous and comparable hardware; handling a reproducible and ideally efficient software stack; managing storage and their access-control lists; rendering results dashboards and summarizing benchmarking results; version controlling code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results. Benchmarking systems are in charge of orchestrating both workflow execution and these benchmarking-specific tasks.

@@

Benchmarks are often focused on finding the `top' performing method, or on generating ranked lists of methods. While this is valid in a few fields with clearly defined goals, such as image recognition \cite{Khan2018-ot} or protein structure prediction \cite{Jumper2021-oq}, bioinformatics tasks are typically evaluated by multiple metrics, some of which are not well understood \cite{Lutge2021-mt, Reinke2024-dw}. Additionally, ground truths are not always well-defined (e.g., derived with uncertainty from other experimental data), multiple tasks (e.g., represented by different reference datasets) may be evaluated for a single method and several ways to assess performance are possible. Even when accompanied by dataset- and/or metric-specific performance measures, scaling and aggregation of results makes it difficult to unpack fine-grained nuances of performance from a final summary table reported in most benchmarks (e.g., \footnote{\url{https://www.nature.com/articles/s41592-021-01336-8/figures/11}}). Strobl \textit{et al.} push back against the `one method fits all datasets’ philosophy and advocate for highlighting dataset or parameter properties that are predictive of method performance \cite{Jelizarow2010-kr, Boulesteix2010-zj, Strobl2024-lp}. This aligns with what a data analyst wants: to identify the best performers in settings most similar to their problem at hand. This requires not only a thorough benchmark design, e.g., tracking of meaningful dataset metadata \cite{Strobl2024-lp}, but also can be supported by a benchmarking system that facilitates flexible navigation of results (e.g., via an interactive dashboard).

@@

Altogether, evaluating parameters as part of a benchmark can give useful insights, but should be linked to a distinct and independent research objective.
For continuous and collaborative benchmarking, parameter tuning and usage should be governed and coordinated between all contributors. Ideally, the benchmarker specifies the parameter scope and related research objectives at the beginning of the benchmark to ensure comparability and interpretability of the results.

@@

In designing a benchmarking system, it may be helpful to specify whether the benchmark is intended to be run `centralized' or `decentralized'. These terms can take different meanings depending on the context; we will focus on the organization of the benchmark infrastructure. In most systems, the code to execute will be tracked in a reference git repository (e.g., on GitHub, GitLab). The workflow execution and the storage of derived datasets (as various stages of results) are typically located either in local deployments (e.g., OEB-VRE for \texttt{OpenEBench}), commercial clouds (e.g., AWS for \texttt{OpenProblems}), commercial continuous integration and continuous delivery platforms (e.g., GitHub actions for \texttt{ncbench}) or could be flexible, e.g., on arbitrary infrastructure (i.e., could be a laptop, server or cloud).

@@

Arguably, most benchmarks could naturally start out as decentralized, in part as a means to properly define the concept and problem that one wants to work on, but also as a way to generate preliminary results and ideas to solidify the design before `going public' (centralization). Thus, users need a system that can build benchmark components in both a centralized and decentralized fashion, i.e., starting decentralized and promoting a benchmark to centralized when it becomes sufficiently mature. Accordingly, the process of centralization is not cost-free: it implies gatekeeping (so method contributors can authenticate to the system and be given access to resources) and adding measures to build trust and transparency.

@@

A related challenge concerns which datasets and metrics to include in a benchmark. On the one hand, diverse sets of datasets and metrics are important to ensure generalizability and avoid overfitting. On the other hand, inclusion of low-quality or redundant datasets or metrics does not benefit the benchmark, and may provide misleading (aggregate) results. Possible mitigations may include sensitivity analyses, evaluating the impact of single datasets or metrics on the overall, aggregated performance of a method, or reducing redundancy by downweighting similar metrics to avoid a specific aspect dominating the evaluation. From a wider perspective, these questions are all related to the governance of a benchmark, a topic that has not received much attention. Ultimately benchmarks may want to, depending on their scale and complexity, implement different governance models. In one model, the benchmarkers have the ultimate authority to decide on what gets included in the benchmark, simplifying the decision making but creating a risk of gatekeeping. In another model, decisions are made collectively by a group of contributors, which may mitigate the risk of gatekeeping at least partially, but creates a more complex decision structure and introduces the need for a conflict resolution strategy. The governance model of a benchmark could also cover aspects related to computing infrastructure and storage, as well as guidelines to manage authorship of any resulting publications or resources. Having the governance model spelled out when launching a benchmark is likely to reduce friction in later stages of the project. In all cases, an advantage of doing things in public is that contributions are logged.

Neutrality of a benchmark is a highly desirable property and it means that efforts have been taken to ensure impartiality or unbiasedness (or at least transparency) of the various choices made (e.g., what datasets to include, how exactly to run methods, how to evaluate methods). Although highly desirable, neutrality is nonetheless extremely tough to assess given only the downstream artifacts of a benchmark; this is an area of governance that deserves further attention to increase trust. One could claim that all MDP benchmarks are non-neutral, since the goal is generally to highlight the virtues of a method. Similarly, ground truth generation and simulations in MDP benchmarks can unconsciously mirror the method capabilities, mechanisms, or strengths, thus biasing evaluation. Furthermore, there are many field experts who may wish to conduct a benchmark (or participate in a collaborative benchmarks) on analyses that involve their own tools, where neutrality is then hard to establish (e.g., perceived conflict of interest to evaluate one's own method) or where deeper knowledge of their own tools might favor the implementation of the latter compared to other competitors. Here, some recent work to mitigate this conflict involves pre-registration, which typically involves specifying the parameters of your study in advance and posting them on a registry \cite{Sullivan2019-wh, Olevska2021-if}; initial formalization of a benchmark could semi-automatically provide the basis for the formal preregistration. In contrast, BOP benchmarks give, in principle, a higher degree of neutrality, because the goal is not simply to advertise the virtues of a single new method; nonetheless, the challenge to assess neutrality directly remains difficult. Other benchmarking models are possible, from challenges to hackathons and here, there is hope that a consensus of the masses can help to establish neutrality. One potential strategy, in the context of the systematic design of benchmarks, is establishing a system where a benchmark definition is parsed into a pre-registration template. Templates exist for simulation studies that could be adapted more generally for benchmarking projects \cite{Siepe2023-tz}. How to reconcile continuity and pre-registration of benchmarks needs further attention.

@@

Software plays an important role in the current scientific landscape \cite{Howison2015-rc}, since it is widely used throughout the scientific process, including data collection, simulation, analysis and reporting. Researchers are encouraged to publish their data and code to enhance transparency and reproducibility in their work. Ideally, this would increase the adoption rate by the wider scientific community. In practice, however, it is sometimes easier to develop new code than to reuse existing software \cite{Trisovic2022-ol}. This leads to the phenomenon of `academic abandonware', where projects are forgotten in code repositories, exacerbating the reproducibility crisis in research; one of the noted challenges in benchmarking was the lack of extensible frameworks \cite{Sonrel2023-te}. A common reason for abandoned projects is their failure to adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) principles \cite{Wilkinson2016-bh}. Another common cause is the life cycle inherent to academic practice: after delivering a research output, there are often no resources to maintain a software package in the longer term; left alone, the likelihood of a particular library compiling or being able to run after 5-10 years will drastically decrease.

@@

The UNIX design philosophy \cite{Pike1984-we} is worth embracing to produce benchmarking systems software that is lightweight, robust, and easily maintainable; ease of use and extensibility are also desirable. Furthermore, to reduce the entry barrier for contributors, the benchmarking system should be capable of running locally with minimal setup. This approach ensures that the system remains relevant and functional even in the absence of centralization. Furthermore, intuitive APIs and sensible defaults \cite{Proctor2018-do} are factors that contribute to highly usable systems. Structured approaches to documentation, including adopting frameworks like diataxis \cite{ProcidaUnknown-vf}, can facilitate adoption and new contributions to a benchmark.

@@

\item Contributions related to individual benchmark components (e.g., datasets, methods, metrics) are expected to be small, typically short scripts that process data and wrap functionality by importing external libraries (where the component implementation itself is developed). Even so, imposing good practices at this level (output validation, testing for abnormal termination, ability to run with a subset of input data) can be beneficial to increase the quality and maintainability of the contributions. The toolset can also enforce validation of metadata annotations (authorship, versioning, licences) for each contributed module.

\item The software dependencies have easily the biggest impact on replicability and maintainability; archiving all combinations of a large dependency tree, across an arbitrary number of base OS images, will exponentially increase retention and maintenance costs. Turning the long-term replicability of a benchmark into a tractable problem is linked to choosing a suitable software management system. Several software management systems have emerged to address these problems, from containerization (e.g., apptainer or singularity) \cite{Kurtzer2017-mn} and automating reproducible and efficient software management, from \texttt{EasyBuild} \cite{Hoste2012-gg} and \texttt{Spack} \cite{Gamblin2015-ll} to distributing software installations such as CernVM-FS \cite{Blomer2013-jk} and combinations of the two such as \texttt{computecanada} \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{Droge2023-ax}.

@@

Even the simplest benchmark designed and run by a single person aiming to compare their method against various `gold standard' methods incorporates the work of many (dataset generators, method developers, operating systems with system dependencies and their developers). Benchmark artifacts are generated by running a workflow \cite{Jackson2021-ke}. Hence, data, code and operating system dependencies are all accessible during the benchmarking process. For reproducibility and transparency and to build trust \cite{Laine2007-py}, this freedom should be extended for others to run (re-run), copy (including fork), distribute (e.g., snapshot), study (e.g., read how another person is implementing my own method), change (bug fix, contribute) and improve (extend) benchmarks. Such freedoms are granted (or restricted) by software and content licenses, as are copyright and authorship and attribution \cite{Kreutzer2014-ua}. Without explicit licenses, even publicly available code is copyrighted and reuse is restricted (to reproduce, distribute, or create derivative works) \cite{Kreutzer2014-ua}.

In practical terms, a benchmarking system needs to be aware of software and data licenses on import, track further changes and attribute them to the (academic) benchmark contributor. A concise summary of license rights and restrictions, and compatibility across commonly-used open-source licenses can be found in Table~\ref{tab:licenses}. Benchmarking components should be contributed with a clear FOSS (free and open source software) license, perhaps by assigning the copyright holder to contributors describing their name, email and  ORCID (as the canonical scholarly identifier), hence allowing seamless reuse, transparency and easing attribution.

@@

Similarly, even if all benchmark code were free software, running them could be hindered by (non-free) proprietary interpreters (e.g., MATLAB). Benchmarking systems should be explicit in allowing or disallowing proprietary software requirements, and be aware and track EULAs (end user license agreements). This conundrum extends to the operating system layer and to system dependencies (e.g., compilers). In terms of benchmark re-runnability, fully FOSS system dependencies arguably offer the highest reusability, and can be handled in a reproducible manner across operating systems and system architectures (e.g., amd64, arm64) \cite{Droge2023-ax}.

@@

\section*{There are various design tradeoffs}

@@

\begin{itemize}
    \item Flexibility vs. complexity: unrestricted freedom for method contributors can mean displacing the burden of maintainability towards the benchmarker. Fewer restrictions (e.g., simpler entry points) means that more methods can be included in a single benchmark; but, it can also imply a higher complexity for the initial setup by the benchmarker. Constraints can also enable validation, which can greatly help with the development process and quality assurance.
    \item  Perfect software reproducibility is possible \cite{Lamb2021-tb}, but it comes with a high engineering cost. It is thus reasonable for a benchmarker to state the desired degree of replicability and the corresponding constraints; different tasks might demand different levels of compliance. 
    \item  Security concerns: decentralized runs lower the participation barrier, but it also increases the potential for attacks to their computing environment. Sandboxed environments (restricting network access, filesystem access etc) can be provided as mitigations. Additionally, static and dynamic code analysis in the continuous integration can help approve potentially risky submissions.
\end{itemize}

@@

Benchmarking involves applying methods to data. The types of data can be diverse and even a single dataset can be stored in multiple formats. A broad distinction can be made according to general vs. specific formats and open vs. proprietary formats. Programming language-specific file formats should be avoided since those formats may change over time, security risks can arise \cite{Huynh2023-pq, Bleih2024-lv}, and in general they are not interoperable. Likewise, proprietary formats should be avoided, since they may change and could require specific (proprietary) software to read. General language-agnostic formats often do not change over time and are thus more interoperable such that these datasets can be made FAIR \cite{Wilkinson2016-bh}. Generally, using established standards is recommended (e.g., SAM, BED, FASTQ, etc for genomics), since related methods are developed around these (typically storage-efficient) formats.

@@

Additionally, which level and scope of data validation needs to be decided. Validations include content integrity (e.g., md5sum), data syntax (e.g., JSON Schema), and data semantics (e.g., SHACL for linked data). Intermediate data and results should be made FAIR, leading to additional decisions, including the scope of data sharing (e.g., only the most important subset), the handling of the metadata (automatic versus manual annotation) and technical aspects (where and how to make available, e.g., GEO or Zenodo). Automatic generation of metadata might require predefined standards and an ontology to fully track how the data was generated \cite{LeboUnknown-om}.

@@

Benchmarking is a cornerstone in bioinformatics, impacting how methods are conceived, developed, published and chosen. Benchmarking is driven by the effort of communities of people with the expertise to address similar (biological and technical) problems. This expertise is leveraged by current practices of community benchmarking, so scientific advance is eased by joint efforts to catalog open problems, generate data to depict (or simulate) them, develop methods to address them, and evaluate their performances. Community benchmarking has been proved successful and very productive \cite{Moult2005-ne,Capella-Gutierrez2017-dh,Luecken2024-fk}.

Here we discuss the motivation and process of community benchmarking at several levels, from definition to execution, while highlighting technologies to potentially ease the technical, scientific and contribution challenges.

@@

MDR acknowledges funding from the Swiss National Science Foundation (grants 200021\_212940 and 310030\_204869) as well as support from swissuniversities P5 Phase B funding (project 23-36_14). CS is supported by the Novartis Research Foundation.

>>>>

\begin{abstract}
Benchmarking, which involves collecting reference datasets and demonstrating method performances, is a requirement for the development of new computational tools, but also becomes a domain of its own to achieve neutral comparisons of methods. Although a lot has been written about how to design and conduct benchmark studies, this Perspective sheds light on a wish list for a computational platform to orchestrate benchmark studies. We discuss various ideas for organizing reproducible software environments, formally defining benchmarks, orchestrating standardized workflows, and how they interface with computing infrastructure.
\end{abstract}

@@

In this work, we give our perspective on how to build a high-standard benchmarking ecosystem in the field of bioinformatics, coming primarily from the viewpoint of methodologists (i.e., researchers with computational background working on tools for data). In particular, our case is one of computational scientists embedded in data-rich biology (i.e., computational biology and bioinformatics), where one of the distinguishing factors is rapid technology development (e.g., single-cell assays coupled to high-throughput sequencing, imaging, etc.), which in turn results in a rapid response from the computational community to develop tools, models, and software for the analysis of such emerging datasets. Although our viewpoint is `bioinformatics', there are analogs in adjacent disciplines where methodologists are creating computational tools in a certain scientific context. One critical step in method development is benchmarking (further defined below), whereby either new methods are compared against existing ones, or a set of existing methods are compared in a more neutral way; for these, we will make use of the terms (and abbreviations) `methods-development papers' (MDPs) and `benchmark-only papers' (BOPs), respectively, defined in a recent meta-analysis study \cite{Cao2023-jz}. Notably, there are various ways to structure the running of a benchmarking study: it can be run by a single person or a small group on a laptop (or a server), or it can be organized as a `challenge' or within the duration of a hackathon (groups of people that come together to competitively assess their approaches and/or populate benchmark components).

Our goal here is to lay out general benchmark parameters and definitions, while abstracting towards an ideal \textit{benchmarking system}. This ranges from general benchmark conceptualization to various technical specifications. We will offer some remarks on the philosophy around the benchmarking activity (e.g., gatekeeping), as well as some technical recommendations (e.g., standards).

@@

Benchmark artifacts can be seen as a collection of data files and (perhaps diverse) source code implementations with a pipeline topology (e.g., methods are run on some input data, and method outputs act as input for metrics) that were ideally run as a workflow in some computing environment. Current practices in single-cell RNA-seq benchmarking show that code is frequently open, reusable and versioned; input datasets are also typically available \cite{Sonrel2023-te}. However, the \textit{extensibility} of benchmarking studies is generally very low, as well as the proportion of benchmarks using a formal workflow system. Thus, a formal system to organize components and orchestrate the crossing of datasets, methods, and metrics would largely impact the current landscape.

@@

Benchmarking as a process deals with tasks other than workflow definition and execution, including: collecting and keeping track of contributions from users; provisioning or being aware of homogeneous and comparable hardware; handling a reproducible and efficient software stack; managing storage and their access-control lists; rendering results dashboards and summarizing results; versioning code, workflow runs, and files; and providing extensive documentation from defining scope or contribution guidelines, to transparent logging and crediting contributions and disseminating results. Benchmarking systems are in charge of orchestrating both workflow execution and these benchmarking-specific tasks.

@@

Benchmarks are often focused on finding the `top' performing method, or on generating ranked lists of methods. While this is valid in a few fields with clearly defined goals, such as image recognition \cite{Khan2018-ot} or protein structure prediction \cite{Jumper2021-oq}, bioinformatics tasks are typically evaluated by multiple metrics, some of which are not well understood \cite{Lutge2021-mt, Reinke2024-dw}. Additionally, ground truths are not always well-defined (e.g., derived with uncertainty from other experimental data), multiple tasks (e.g., represented by different reference datasets) may be evaluated for a single method and several ways to assess performance are possible. Even when accompanied by dataset- and/or metric-specific performance measures, scaling and aggregation of results make it difficult to unpack fine-grained nuances of performance from a final summary table reported in most benchmarks (e.g., \footnote{\url{https://www.nature.com/articles/s41592-021-01336-8/figures/11}}). Strobl \textit{et al.} push back against the `one method fits all datasets’ philosophy and advocate for highlighting dataset or parameter properties that are predictive of method performance \cite{Jelizarow2010-kr, Boulesteix2010-zj, Strobl2024-lp}. This aligns with what a data analyst wants: to identify the best performers in settings most similar to their problem at hand. This requires not only a thorough benchmark design, e.g., tracking of meaningful dataset metadata \cite{Strobl2024-lp}, but also can be supported by a benchmarking system that facilitates flexible navigation of results (e.g., via an interactive dashboard).

@@

Altogether, evaluating parameters as part of a benchmark can give useful insights, but should be linked to a distinct and independent research objective.

For continuous and collaborative benchmarking, parameter tuning and usage should be governed and coordinated between all contributors. Ideally, the benchmarker specifies the parameter scope and related research objectives at the beginning of the benchmark to ensure comparability and interpretability of the results.

@@

In designing a benchmarking system, it may be helpful to specify whether the benchmark is intended to be run `centralized' or `decentralized'. These terms can take different meanings depending on the context, but here the benchmark infrastructure location is intended. In most systems, the code to execute will be tracked in a reference \texttt{git} repository (e.g., on GitHub, GitLab). The workflow execution and the storage of derived datasets, and various stages of results are typically located either in local deployments (e.g., OEB-VRE for \texttt{OpenEBench}), commercial clouds (e.g., AWS for \texttt{OpenProblems}), commercial continuous integration and continuous delivery platforms (e.g., GitHub actions for \texttt{ncbench}) or could be flexible, e.g., on arbitrary infrastructure (i.e., could be a laptop, server or cloud).

@@

Arguably, most benchmarks could naturally start out as decentralized, in part as a means to properly define the concept and problem that one wants to work on, but also as a way to generate preliminary results and ideas to solidify the design before `going public' (centralization). Thus, users need a system that can build benchmark components in both a centralized and decentralized fashion, i.e., starting decentralized and promoting a benchmark to centralized when it becomes sufficiently mature. Accordingly, the process of centralization is not cost-free: it implies gatekeeping (so method contributors can authenticate to the system and be given access to computational resources) and adding measures to build trust and transparency.

@@

A related challenge concerns which datasets and metrics to include in a benchmark. On the one hand, diverse sets of datasets and metrics are important to ensure generalizability and avoid overfitting. On the other hand, inclusion of low-quality or redundant datasets or metrics does not benefit the benchmark, and may provide misleading (aggregate) results. Possible mitigations may include sensitivity analyses, evaluating the impact of single datasets or metrics on the overall, aggregated performance of a method, or reducing redundancy by downweighting similar metrics to avoid a specific aspect dominating the evaluation. From a wider perspective, these questions are all related to the governance of a benchmark, a topic that has not received much attention. Ultimately benchmarks may want to, depending on their scale and complexity, implement different governance models. In one model, the benchmarkers have the ultimate authority to decide on what gets included in the benchmark, simplifying the decision making but creating a risk of gatekeeping. In another model, decisions are made collectively by a group of contributors, which may mitigate the risk of gatekeeping at least partially, but creates a more complex decision structure and introduces the need for a conflict resolution strategy. The governance model of a benchmark could also cover aspects related to computing infrastructure and storage, as well as guidelines to manage authorship of any resulting publications or resources. Having the governance model spelled out when launching a benchmark is likely to reduce friction at later stages of the project. In all cases, an advantage of doing things in public is that contributions are logged.

Neutrality is a highly desirable property in a benchmark: it means that efforts have been taken to ensure impartiality or unbiasedness (or at least transparency) of the various choices made (e.g., what datasets to include, how exactly to run methods, how to evaluate methods). Although highly desirable, neutrality is nonetheless extremely tough to assess given only the downstream artifacts of a benchmark; this is an area of governance that deserves further attention to increase trust. One could claim that all MDP benchmarks are non-neutral, since the goal is generally to highlight the virtues of a method. Similarly, ground truth generation and simulations in MDP benchmarks can unconsciously mirror the method capabilities, mechanisms, or strengths, thus biasing evaluation. Furthermore, there are many field experts who may wish to conduct a benchmark (or participate in a collaborative benchmark) on analyses that involve their own tools, where neutrality is then hard to establish (e.g., a perceived conflict of interest to evaluate one's own method) or where a deeper knowledge of their own tools might favor the implementation of the latter compared to other competitors. Here, some recent work to mitigate this conflict involves pre-registration, which typically involves specifying the parameters of your study in advance and posting them on a registry \cite{Sullivan2019-wh, Olevska2021-if}; initial formalization of a benchmark could semi-automatically provide the basis for the formal preregistration. In contrast, BOP benchmarks give, in principle, a higher degree of neutrality, because the goal is not simply to advertise the virtues of a single new method; nonetheless, the challenge to assess neutrality directly remains difficult. Other benchmarking models are possible, from challenges to hackathons and here, there is hope that a consensus of the masses can help to establish neutrality. One potential strategy, in the context of the systematic design of benchmarks, is establishing a system where a benchmark definition is parsed into a pre-registration template. Templates exist for simulation studies that could be adapted more generally for benchmarking projects \cite{Siepe2023-tz}. How to reconcile continuity and pre-registration of benchmarks needs further attention.

@@

Software plays an important role in the current scientific landscape \cite{Howison2015-rc}, since it is widely used throughout the scientific process, including data collection, simulation, analysis and reporting. Researchers are encouraged to publish their data and code to enhance transparency and reproducibility in their work. Ideally, this would increase benchmark (software) reuse by the wider scientific community. In practice, however, it is sometimes easier to develop new code than to reuse existing software \cite{Trisovic2022-ol}. This leads to the phenomenon of `academic abandonware', where projects are forgotten in code repositories, exacerbating the reproducibility crisis in research; one of the noted challenges in benchmarking was the lack of extensible frameworks \cite{Sonrel2023-te}. A common reason for abandoned projects is their failure to adhere to the FAIR (Findable, Accessible, Interoperable, Reusable) principles \cite{Wilkinson2016-bh}. Another common cause is the life cycle inherent to academic practice: after delivering a research output, there are often no resources to maintain a software package in the longer term; left alone, the likelihood of a particular library compiling or being able to run after 5-10 years will drastically decrease.

@@

The UNIX design philosophy \cite{Pike1984-we} is worth embracing to produce software for benchmarking systems that are lightweight, robust, and easily maintainable; ease of use and extensibility are also desirable. Furthermore, to reduce the entry barrier for contributors, the benchmarking system should be capable of running locally with minimal setup. This approach ensures that the system remains relevant and functional even in the absence of centralization. Furthermore, intuitive APIs and sensible defaults \cite{Proctor2018-do} are factors that contribute to highly usable systems. Structured approaches to documentation, including adopting frameworks like diataxis \cite{ProcidaUnknown-vf}, can facilitate adoption and new contributions to a benchmark.

@@

\item Contributions related to individual benchmark components (e.g., datasets, methods, metrics) are expected to be small, typically short scripts that process data and wrap functionality by importing external libraries (where the component implementation itself is developed). Even so, imposing good practices at this level (output validation, testing for abnormal termination, ability to run with a subset of input data) can be beneficial to increase the quality and maintainability of the contributions. The toolset can also enforce the validation of metadata annotations (authorship, versioning, licences) for each contributed module.

\item The software dependencies have easily the biggest impact on replicability and maintainability; archiving all combinations of a large dependency tree, across an arbitrary number of base OS images, will exponentially increase retention and maintenance costs. Turning the long-term replicability of a benchmark into a tractable problem is linked to choosing a suitable software management system. Several software management systems have emerged to address these problems, from containerization (e.g., apptainer or singularity) \cite{Kurtzer2017-mn} and automation of reproducible and efficient software management (\texttt{EasyBuild} \cite{Hoste2012-gg} or \texttt{Spack} \cite{Gamblin2015-ll}) to distributing software installations such as CernVM-FS \cite{Blomer2013-jk},  or combinations of the two such as \texttt{computecanada} \cite{Boissonneault2019-wm} and \texttt{EESSI} \cite{Droge2023-ax}.

@@

Even the simplest benchmark designed and run by a single person aiming to compare their method against various `gold standard' methods incorporates the work of many (dataset generators, method developers, operating systems with system dependencies and their developers). Benchmark artifacts are generated by running a workflow \cite{Jackson2021-ke}. Hence, data, code and operating system dependencies are all accessible during the benchmarking process. For reproducibility and transparency and to build trust \cite{Laine2007-py}, this freedom should be extended for others to run (re-run), copy (including fork), distribute (e.g., snapshot), study (e.g., read how another person is implementing my own method), change (bug fix, contribute) and improve (extend) benchmarks. Such freedoms are granted (or restricted) by software and content licenses, as are copyright and authorship and attribution \cite{Kreutzer2014-ua}. Without explicit licenses, even publicly available code is copyrighted by default, and reuse is restricted (to reproduce, distribute, or create derivative works) \cite{Kreutzer2014-ua}.

In practical terms, a benchmarking system needs to be aware of software and data licenses, track further changes and attribute them to the (academic) benchmark contributor. A concise summary of license rights and restrictions, and compatibility across commonly-used open-source licenses can be found in Table~\ref{tab:licenses}. Benchmark contributions should bring a clear FOSS (free and open source software) license, perhaps by assigning the copyright holder to contributors describing their name, email and  ORCID (as the canonical scholarly identifier), hence allowing seamless reuse, transparency and easing attribution.

@@

Similarly, even if all benchmark code were free software, running it could be hindered by (non-free) proprietary interpreters (e.g., MATLAB). Benchmarking systems should be explicit in allowing or disallowing proprietary software requirements, and be aware and track EULAs (end user license agreements). This conundrum extends to the operating system layer and to system dependencies (e.g., compilers). In terms of benchmark re-runnability, fully FOSS system dependencies arguably offer the highest reusability, and can be handled in a reproducible manner across operating systems and system architectures (e.g., amd64, arm64) \cite{Droge2023-ax}.

@@

\section*{There are various hidden design tradeoffs}

@@

\begin{itemize}
    \item Flexibility vs. complexity: unrestricted freedom for method contributors can mean displacing the burden of maintainability towards the benchmarker. Fewer restrictions (e.g., simpler entry points) means that more methods can be included in a single benchmark; but, it can also imply a higher complexity for the initial setup by the benchmarker. Constraints can also enable validation, which can greatly help with the development process and quality assurance.
    \item  Perfect software reproducibility is possible \cite{Lamb2021-tb}, but it comes with a high engineering cost. It is thus reasonable for a benchmarker to state the desired degree of replicability and the corresponding constraints; different tasks might demand different levels of compliance. 
    \item  Security concerns: decentralized runs lower the participation barrier, but it also increases the potential for attacks on their computing environment. Sandboxed environments (restricting network access, filesystem access etc) can be provided as mitigation. Additionally, static and dynamic code analysis in the continuous integration process can aid in approving potentially risky submissions.
\end{itemize}

@@

Benchmarking involves applying methods to data. The types of data can be diverse and even a single dataset can be stored in multiple formats. A broad distinction can be made according to general vs. specific formats and open vs. proprietary formats. File formats specific to a programming language should be avoided since those formats may change over time, security risks can arise \cite{Huynh2023-pq, Bleih2024-lv}, and in general they are not interoperable. Likewise, proprietary formats should be avoided, since they may change and could require specific (proprietary) software to read. General language-agnostic formats often do not change over time and are thus more interoperable such that these datasets can be made FAIR \cite{Wilkinson2016-bh}. Generally, using established standards is recommended (e.g., SAM, BED, FASTQ, etc for genomics), since related methods are developed around these (typically storage-efficient) formats.

@@

Additionally, which level and scope of data validation needs to be decided. Validations include content integrity (e.g., md5sum), data syntax (e.g., JSON Schema), and data semantics (e.g., SHACL for linked data). Intermediate data and results should be made FAIR, leading to additional decisions, including the scope of data sharing (e.g., only the most important subset), the handling of the metadata (automatic versus manual annotation) and the provider (where and how to make available, e.g., GEO or Zenodo). Automatic generation of metadata might require predefined standards and an ontology to fully track how the data was generated \cite{LeboUnknown-om}.

@@

Benchmarking is a cornerstone practice in bioinformatics, impacting how methods are conceived, developed, published and chosen. Benchmarking is driven by the effort of people with the expertise to address (biological and technical) problems. This expertise can be leveraged by current practices of community benchmarking, so scientific advance is eased by joint efforts to catalog open problems, generate (or simulate) appropriate input data, develop methods to address them, and evaluate their performances. Community benchmarking has been proved successful and very productive \cite{Moult2005-ne,Capella-Gutierrez2017-dh,Luecken2024-fk}.

Here we discussed the motivation and process of community benchmarking at several levels, from definition to execution, while highlighting technologies to potentially ease the technical, scientific and contribution challenges.

@@

MDR acknowledges funding from the Swiss National Science Foundation (grants 200021\_212940 and 310030\_204869) as well as support from swissuniversities P5 Phase B funding (project 23-36\_14). CS is supported by the Novartis Research Foundation.
